---
title: "Assignment7"
author: "Yerin Jung"
date: "7/9/2018"
output:   
  html_document:
    theme: lumen
    number_sections: yes
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


# Data Preparation 

```{r packages, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(plotly)
library(kableExtra)
library(h2o)
library(MeanShift)
library(ggfortify)
library(knitr)
```

```{r echo=FALSE}
setwd("/Users/Jung-yerin/Desktop/SUMMER_2018/AD/Week7")
ccard <- read_csv("ccard_ready.csv")
```

For this assignment, I am going to use credit card data.  
The features here are the same with the ones that I have created for the Assignment 6.  


```{r echo=FALSE}
kable(head(ccard), "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), full_width = FALSE, font_size=10)%>%
  scroll_box(height = "20%", width = "90%")

```

# Autoencoder



In this section,  
  - First, I will plot the data set using one hidden layer. Since Autoencoder uses a bottleneck model, we can do dimension-reduction via Autoencoder.  
  - Second, I will detect anomalies using the mean squared error between the real value (input) and reconstructed value (output).  
  
  
  
  To do the above two agendas, I will initiate an h20 model.  
  
  
```{r}
h2o.init(nthreads=-1,enable_assertions = FALSE)
```


  
   Next, I changed the data frame into an h2o-type data frame.  
   
   
```{r results='hide'}
ccard_hf <- as.h2o(ccard)
```
  

```{r echo=FALSE}
kable(head(ccard_hf), "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), full_width = FALSE, font_size=10)%>%
  scroll_box(height = "20%", width = "90%")
```

  
  Then, I split the train and test model.  
  

```{r}
train_test <- h2o.splitFrame(ccard_hf, 
                         ratios = c(0.4, 0.4), 
                         seed = 101)

train  <- train_test[[1]]
test1  <- train_test[[2]]
test2 <- train_test[[3]]

features <- colnames(train)
```






  Now I am going to build an autoencoder model. 
  I will use 3 hidden layers and will adjust later to see various results depending on the number of units in the hidden layers.  
  
  

```{r results='hide'}
model_nn <- h2o.deeplearning(x = features,
                             training_frame = train,
                             model_id = "model_nn",
                             autoencoder = TRUE,
                             seed = 101,
                             hidden = c(10, 2, 10), 
                             epochs = 100,
                             activation = "Tanh")
```



## Dimensionality Reduction
  
  
  
  The below codes represent the way that I used to perform dimensionality reduction.   
  
  

```{r results='hide' }
train_features <- h2o.deepfeatures(model_nn, train, layer = 2) %>%
  as.data.frame()
```


```{r}

ggplot(train_features, aes(x = DF.L2.C1, y = DF.L2.C2)) +
  geom_point(alpha = 0.1)+
  ggtitle("Low dimensionality with Hidden Layer Units [10, 2, 10]")
```


## Anomaly Detection  

  
  Now, I am going to build a model to detect anomalies in the data set, using `h2o.anomaly` function.  
  
  
```{r}
anomaly <- h2o.anomaly(model_nn, test2) %>%
  as.data.frame() %>%
  tibble::rownames_to_column()
mean_mse <- anomaly %>%
  summarise(mean = mean(Reconstruction.MSE))
```

```{r}
ggplot(anomaly, aes(x = as.numeric(rowname), y = Reconstruction.MSE)) +
  geom_point(alpha = 0.3) +
  geom_hline(data = mean_mse, aes(yintercept = mean), color="blue") +
  annotate(geom="text", x=100, y=0.05, label="Total Mean", color="blue")+
  ggtitle("Reconstruction MSE for instances")
```


As we can see from the below table, $10$ cases were detected as outliers.  


```{r echo=FALSE}
anomaly <- anomaly %>%
  mutate(outlier = ifelse(Reconstruction.MSE > 0.02, "outlier", "no_outlier"))

anomaly_df <- anomaly %>% 
  as.data.frame()

test2_df <- test2 %>%
  as.data.frame()

test2_anomaly <- data.frame( Agency=test2_df$Agency_Name, Reconstruction.MSE=anomaly_df$Reconstruction.MSE, outlier=anomaly_df$outlier)

ano_result <- test2_anomaly %>%
  group_by(outlier) %>%
  summarise(n = n())


kable(ano_result, "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), full_width = FALSE)
```




   To identify the agencies that were detected as outliers, I plotted a histogram.  
  
  
```{r}

test2_anomaly %>%
  filter(outlier=="outlier")%>%
  ggplot(aes(x=Agency))+
  geom_histogram(stat="count", fill="skyblue")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Outliers from h2o")
```


  We can find $8$ agencies from the above histogram.  
  
  
  

# Mean-Shift


  Due to the computational limit, I sampled only 20% of the entire dataset.  
  After sampling, I scaled them.  
  
  
```{r}
set.seed(101)

ind <-sample(nrow(ccard), floor(nrow(ccard)*0.2))
ccard_sample <- ccard[ind,]
ccard_sd<-scale(ccard_sample[,3:26])
ccard_data <-t(ccard_sd) 
dim(ccard_data)
```

## h=10.0

 First, I will try msClustering with `h`=$10$.  
 I will gradually increase the size of the bandwidth (10, 15, and 20) and compared the results.  
 
 
 
  
```{r results='hide'}
ms_clustering1 <- msClustering( ccard_data, h=10.0 )
```


```{r}
# Plotting using PCA
ccard_ms1 <- ccard_sample
ccard_ms1[,'ms_label'] <- ms_clustering1[2]
ccard_ms1[,'ms_label'] <- as.factor(ccard_ms1$ms_label)
autoplot(prcomp(ccard_ms1[,3:25]), data = ccard_ms1, colour = 'ms_label',frame=TRUE)
```

 
  As we can see from the below table, most of the values are labeled as one, and there were 25 labels in total.  
  
```{r}
# Counting numbers by label
anomaly_summary <- ccard_ms1 %>%
  group_by(ms_label) %>%
  summarise(n_count=n())


kable(anomaly_summary, "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), full_width = FALSE, font_size=10)%>%
  scroll_box(width = "500px", height = "300px")
```

  
  
 
 Finally, I looked at the name of the agencies, who have been detected as outliers (labeled other than 1).  



```{r}
ccard_ms1 %>%
  filter(ms_label != 1) %>%
  ggplot(aes(x=Agency_Name))+
  geom_histogram(stat = "count", fill="skyblue")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Outliers from MeanShift with h=10.0")
```
  

  
  
  
## h=15.0
  
  
  I tried the same approach with the parameter `h`=$15$.  
  
```{r results='hide'}
ms_clustering2 <- msClustering( ccard_data, h=15.0 )
```
 
  
```{r echo=FALSE}

ccard_ms2 <- ccard_sample
ccard_ms2[,'ms_label'] <- ms_clustering2[2]
ccard_ms2[,'ms_label'] <- as.factor(ccard_ms2$ms_label)
autoplot(prcomp(ccard_ms2[,3:25]), data = ccard_ms2, colour = 'ms_label',frame=TRUE)
```
  
  
  As we can see from the below table, with the parameter `h`=$15$, fewer instances were detected than when `h`=$10$ as outliers (other than 1).    
  
  
```{r echo=FALSE}
anomaly_summary2 <- ccard_ms2 %>%
  group_by(ms_label) %>%
  summarise(n_count=n())

kable(anomaly_summary2, "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), full_width = FALSE, font_size=10)%>%
  scroll_box(width = "500px", height = "300px")
```
  
  
  Finally, I also looked at the name of the agencies, who have been detected as outliers (labeled other than 1). 
  
  
```{r echo=FALSE}
ccard_ms2 %>%
  filter(ms_label != 1) %>%
  ggplot(aes(x=Agency_Name))+
  geom_histogram(stat = "count", fill="skyblue")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Outliers from MeanShift with h=15.0")
```
  
  In this case, only `OKLAHOMA STATE UNIVERSITY` and `OKLAHOMA MILITARY DEPARTMENT` have more than 1 outliers.  
  
  

## h=20.0 
  
  
  For the last parameter, I tried the same approach with the parameter `h`=$20$.  
  
```{r results='hide'}
ms_clustering3 <- msClustering( ccard_data, h=20.0 )
```


  
```{r echo=FALSE}
ccard_ms3 <- ccard_sample
ccard_ms3[,'ms_label'] <- ms_clustering3[2]
ccard_ms3[,'ms_label'] <- as.factor(ccard_ms3$ms_label)
autoplot(prcomp(ccard_ms3[,3:25]), data = ccard_ms3, colour = 'ms_label',frame=TRUE)
```

  
  As we can see from the below table, with the parameter `h`=$20$, even fewer instances were detected as outliers (other than 1).  
  Only 10 were detected as outliers.  
  
  
```{r echo=FALSE}
anomaly_summary3<- ccard_ms3 %>%
  group_by(ms_label) %>%
  summarise(n_count=n())

kable(anomaly_summary3, "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), full_width = FALSE, font_size=10)%>%
  scroll_box(height = "20%", width = "90%")
```
  
  
  If we plot the name of the agencies, we can see 10 different agencies.  
  
  

```{r echo=FALSE}
ccard_ms3 %>%
  filter(ms_label != 1) %>%
  ggplot(aes(x=Agency_Name))+
  geom_histogram(stat = "count", fill="skyblue")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Outliers from MeanShift with h=20.0")
```


# Autoencoder and Meanshift
  
  Finally, I am going to compare the name of agencies who were detected as anomalies from `MeanShift` and `Autoencoder`.  
  I chose `h`=$15$ for `MeanShift`.  
    
    
    
  As we can see from the below table, there were several agencies who appeared more than once either in each method only or both methods respectively.    
  
  
```{r}
ms_agency <- ccard_ms2 %>%
  filter(ms_label != 1) %>%
  dplyr::select(Agency_Name)
names(ms_agency)[1] <-"Agency"

auto_agency <-test2_anomaly %>%
  filter(outlier=="outlier") %>%
  dplyr::select(Agency)

df_combined <- rbind(ms_agency, auto_agency)

final <- df_combined %>% group_by(Agency)%>%
  summarize(count_agency=n())

kable(final, "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), full_width = FALSE, font_size=10) %>%
  row_spec(c(6, 10, 15, 16, 19), bold = T, color = "white", background = "skyblue")%>%
  scroll_box(width = "500px", height = "300px")
```










