---
title: "Assignment 4"
author: "Yerin Jung"
date: "6/15/2018"
output:   
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

#########Set Up###########

```{r}
source("health-functions.R")
```


```{r}
#setwd("/Users/Jung-yerin/Desktop/SUMMER_2018/AD/AnomalyDetection/week4")
payment <- read_csv("inpatientCharges.csv")
```

```{r}
# Changing column names to remove space
names(payment) <- c("DRG_Def", "Pro_Id", "Pro_Name","Pro_Add","Pro_City","Pro_State","Pro_Zip",
                    "HRR","Tot_Discharges","Avg_Cov_Charges","Avg_Tot_Payments","Avg_Med_Payments")
summary(payment)
```

```{r}
# Parsing the numbers from Drug definition to see graphs effectively 
payment <- payment %>%
  mutate(DRG_Num=substr(x = DRG_Def, start=1, stop=3))
```


# Comparing Standardized Total Payment with The Other Providers 


## Lead
  
  Generally, total payment of a care depends on the region(state) and type of drug function.  
As such, to see whether there is abnormally high total payment (`Avg_Tot_Payments`), I decided to standardize the value, using the mean and standard deviation of `Avg_Tot_Payments` grouped by Drug Definition (`Drug_Num`) and Provider State (`Pro_State`).  

  To do so, first, I should count the number of providers in each drug definition group. If there are just a few numbers of providers in the group, it might be less rational to compare the total payments. (In many cases there was only one provider in a drug function group.)    
  
  
  
  
## EDA


```{r}
# Making groups dependin on the number of providers in each drug group
payment1 <- payment %>% 
  group_by(DRG_Def) %>%
  mutate(n_row_group=n())%>%
  ungroup()%>%
  mutate(group_class= cut(n_row_group, breaks = 3, labels = c("1_to_945", "946_to_1890", "1890_to_2837")))

payment1 %>%
  select(DRG_Def,group_class) %>%
  tail()

# The graph of the third group, which has relatively many providers in each drug group
payment1 %>%
  filter(group_class=="1890_to_2837")%>%
  ggplot(aes(x=Avg_Tot_Payments, y=DRG_Num))+
  geom_point(colour="skyblue")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  scale_y_discrete(label=abbreviate)

```

  
  
  
  Here, as I mentioned, I divided the dataset into three groups: `1_to_945`, `946_to_1890`, and `1890_to_2837`:  
  - `1_to_945`: The number of entities in a drug definition group (`DRG_Def`) is a range of 1 to 946   
  - `946_to_1890`: The number of entities in a drug definition group (`DRG_Def`) is a range of 946 to 1890  
  - `1890_to_2837`: The number of entities in a drug definition group (`DRG_Def`) is a range of 1890 to 2837  
  
  
  Then, I plotted `Avg_Tot_Payments` from only `1890_to_2837` group, to see the general trends.  
  It looks like we see some outliers, though, we should not be so sure, since the price can also vary from states.  
  
  
  
```{r}
# Making Three Columns after grouping by DRG_Num and Pro_State
payment1<-payment1 %>%
  group_by(DRG_Num, Pro_State) %>%
  mutate(Mu_DRG_State_Total_Payment=mean(Avg_Tot_Payments), 
         Sd_DRG_State_Total_Payment=sd(Avg_Tot_Payments),
         Std_DRG_State_Total_Payment=(Avg_Tot_Payments-Mu_DRG_State_Total_Payment)/Sd_DRG_State_Total_Payment) %>%
  ungroup()

# Viewing some examples
payment1 %>%
  select(DRG_Num,Pro_State,Mu_DRG_State_Total_Payment,Sd_DRG_State_Total_Payment,Std_DRG_State_Total_Payment) %>%
  filter(DRG_Num=="310")%>%
  head()

# Plotting standardized total_payment
payment1 %>%
  filter(group_class=="1890_to_2837"|group_class=="946_to_1890" ) %>%
  ggplot(aes(x=Pro_Id, y=Std_DRG_State_Total_Payment))+
  geom_point(aes(color=Pro_State))+
  theme(axis.text.x = element_blank())

```

  After dividing into three group depending on the number of providers in drug_definition,    
  I filtered only the second and third group, which basically have more than 946 providers in each drug definition group.    
  
  I generated three columns as follows:    
  - `Mu_DRG_State_Total_Payment`: Mean of `Avg_Tot_Payments` grouped by drug group (`Drug_Num`) and state (`Prov_State`).  
  - `Sd_DRG_State_Total_Payment`: Standard deviation of `Avg_Tot_Payments` grouped by drug group (`Drug_Num`) and state (`Prov_State`).  
  - `Std_DRG_State_Total_Payment`: Standardized `Avg_Tot_Payments` using Z-score with `Mu_DRG_State_Total_Payment` and `Sd_DRG_State_Total_Payment`  
  
  
  
$$Z=\frac{X-\mu }{\sigma }=\frac{AvgTotalPayments-Mu(AvgTotalPayments)}{Sd(AvgTotalPayments)}$$
  
  
  


## Conclusion

  As we can see from the above colorful graph, we can notice a few distinctively high standardized values, comparing to the other providers. We might want to further exam those values.  
  Since they were outliers even after they were standardized by the mean grouped by state and drug as well, I believe that it has room to be a fraud case.  
  
  


# Standardized Difference Between Medical Coverage and Medical Payment

## Lead
  
  
  
According to [Data.CMS](https://data.cms.gov/Medicare-Inpatient/National-Summary-of-Inpatient-Charge-Data-by-Medic/efwk-h4x3), `Average Covered Charges` means the average charge of all provider's services covered by Medicare for discharges in the DRG.  
Since this would vary from hospital to hospital, we should not simply assume a fraud, just because a provider overly charged Medicare to pay than they were supposed to cover.  
However, we can calculate the mean difference of `Average Covered Charges` and `Average Medicare Payments` per each Drug group, and see whether an entity has significant difference than the other in the same group.  


## EDA


```{r}
# Making Four New Columns
payment2<-payment1 %>%
  mutate(Diff_MedCov_MedPay=Avg_Cov_Charges-Avg_Med_Payments) %>%
  group_by(DRG_Num) %>%
  mutate(Mu_Diff=mean(Diff_MedCov_MedPay), Sd_Diff=sd(Diff_MedCov_MedPay)) %>%
  mutate(Standard_Diff=(Diff_MedCov_MedPay-Mu_Diff)/Sd_Diff)%>%
  ungroup()

payment2 %>%
  select(DRG_Num, Diff_MedCov_MedPay, Mu_Diff, Sd_Diff, Standard_Diff ) %>%
  head()

# Plotting a histogram
payment2 %>%
  ggplot(aes(x=Standard_Diff))+
  geom_histogram()

# Filtering intresting data
payment2 %>% filter(Standard_Diff < -2)
```
  
  First, I made three columns `Diff_MedCov_MedPay`, `Mu_Diff`, `Sd_Diff`, and `Standard_Diff`.  
  - `Diff_MedCov_MedPay`: The difference between the amount that is covered by Medicare(`Avg_Cov_Charges`) and the amount Medicare actually paid (`Avg_Med_Payments`)  
  - `Mu_Diff`: Average of `Diff_MedCov_MedPay` grouped by drug definition(`Drug_Num`). I grouped by drug definition, since some drugs might just require more money than it is generally covered.  
  - `Sd_Diff`: Standard deviation of `Diff_MedCov_MedPay`.  
  - `Standard_Diff`: Standardized values of `Diff_MedCov_MedPay`, using Z-score formula with `Mu_Diff` and `Sd_Diff`   

$$Z=\frac{X-\mu }{\sigma }=\frac{DiffMedCovMedPay-MuDiff}{Sd_Diff}$$
  
  
  
## Conclusion

  As we can see from the above graph, most of the providers were centered around zero. However, there were several providers with the standardized difference even less than -2.  
  As such, I looked at the providers with the standardized difference even less than -2.  
  As we can see from the above chart, there were $17$ providers who have a standardized difference of less than -2.  
  This means the amount of money that Medicare paid to the providers was significantly higher than the amount to the other providers in the same drug definition.  
  We may assume that those cases have a possibility to be a fraud, and examine them.  
  
  
  

