---
title: "Assignment8"
author: "Yerin Jung"
date: "July 17, 2018"
output: 
  ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r packages, include=FALSE}
setwd("/Users/Jung-yerin/Desktop/SUMMER_2018/AD/Week8")
source("assignment8.R")
```



## Contents

- Data Preparation
- GLM
  - Open_Close_PctChange
  - Daily_Return
- Random Forest
  - Open_Close_PctChange
  - Daily_Return  
- Final Results & Business Insights



```{r dataset, include=FALSE}
setwd("/Users/Jung-yerin/Desktop/SUMMER_2018/AD/Week7")
ccard <- read_csv("ccard_ready.csv")
```

```{r data, include=FALSE}
setwd("/Users/Jung-yerin/Desktop/SUMMER_2018/AD/Week1/Data/Stocks")
sp500URL <- getURL("https://raw.githubusercontent.com/datasets/s-and-p-500-companies/master/data/constituents.csv")
sp500File <- file.path('data', 'constituents.csv')

if (file.exists(sp500File)) {
  sp500URL <- sp500File
}

# read in the SP500 information
sp500Members <- read_csv( sp500URL )

# read in the one file
stockDF <- read_csv("stock_data.csv", col_types = c("Dddddnnc")) %>%
  arrange(Symbol, Date)

stockDF$Symbol <- toupper(stockDF$Symbol)

sp500DF <- stockDF %>% filter (Symbol %in% sp500Members$Symbol)

gStockDF <- sp500DF %>% 
  group_by(Symbol) %>%
  arrange(Date)

gStockDF <- gStockDF[complete.cases(gStockDF),]
```


```{r features, include=FALSE}
# first set of features
df2 <- gStockDF %>% 
  mutate(
    Open_Close_PctChange = (Close-Open)/Open * 100,
    
    Open_Close_Delt = Delt(Open, Close, k = 0) * 100,
    
    Open_Change   = Open   - lag(Open),
    High_Change   = High   - lag(High),
    Low_Change    = Low    - lag(Low),
    Close_Change  = Close  - lag(Close),
    Volume_Change = Volume - lag(Volume),
    
    Daily_Return  = Close / Open,
    
    Open_PctChange   = Open_Change   / lag(Open)   * 100,
    High_PctChange   = High_Change   / lag(High)   * 100,
    Low_PctChange    = Low_Change    / lag(Low)    * 100,
    Close_PctChange  = Close_Change  / lag(Close)  * 100,
    Volume_PctChange = Volume_Change / lag(Volume) * 100,
    
    Open_Mean10   = roll_mean( Open,   10, fill = NA, na.rm = TRUE ),
    High_Mean10   = roll_mean( High,   10, fill = NA, na.rm = TRUE ),
    Low_Mean10    = roll_mean( Low,    10, fill = NA, na.rm = TRUE ),
    Close_Mean10  = roll_mean( Close,  10, fill = NA, na.rm = TRUE ),
    Volume_Mean10 = roll_mean( Volume, 10, fill = NA, na.rm = TRUE ),

    Open_Mean10_R   = roll_meanr( Open,   10, fill = NA, na.rm = TRUE ),
    High_Mean10_R   = roll_meanr( High,   10, fill = NA, na.rm = TRUE ),
    Low_Mean10_R    = roll_meanr( Low,    10, fill = NA, na.rm = TRUE ),
    Close_Mean10_R  = roll_meanr( Close,  10, fill = NA, na.rm = TRUE ),
    Volume_Mean10_R = roll_meanr( Volume, 10, fill = NA, na.rm = TRUE ),
    
    Open_SD30   = roll_sd( Open,   30, fill = NA, na.rm = TRUE ),
    High_SD30   = roll_sd( High,   30, fill = NA, na.rm = TRUE ),
    Low_SD30    = roll_sd( Low,    30, fill = NA, na.rm = TRUE ),
    Close_SD30  = roll_sd( Close,  30, fill = NA, na.rm = TRUE ),
    Volume_SD30 = roll_sd( Volume, 30, fill = NA, na.rm = TRUE ),
    
    Open_VAR30   = roll_var( Open,   30, fill = NA, na.rm = TRUE ),
    High_VAR30   = roll_var( High,   30, fill = NA, na.rm = TRUE ),
    Low_VAR30    = roll_var( Low,    30, fill = NA, na.rm = TRUE ),
    Close_VAR30  = roll_var( Close,  30, fill = NA, na.rm = TRUE ),
    Volume_VAR30 = roll_var( Volume, 30, fill = NA, na.rm = TRUE ),
    
    Open_EMA10   = EMA(Open,   n = 10),
    High_EMA10   = EMA(High,   n = 10),
    Low_EMA10    = EMA(Low,    n = 10),
    Close_EMA10  = EMA(Close,  n = 10),
    Volume_EMA10 = EMA(Volume, n = 10),
    
    Open_SMA10   = SMA(Open,   n = 10),
    High_SMA10   = SMA(High,   n = 10),
    Low_SMA10    = SMA(Low,    n = 10),
    Close_SMA10  = SMA(Close,  n = 10),
    Volume_SMA10 = SMA(Volume, n = 10)

  ) %>%
  arrange(Symbol, Date)
```

```{r features2, include=FALSE}
# add SP500 characteristics (sector)
df3 <- df2 %>% left_join(sp500Members, by = "Symbol")

# add weekdays and quarters 
df3$DOW <- weekdays(df3$Date)
df3$Quarter <- quarters(df3$Date)

df3$Year <- as.numeric(format(df3$Date, "%Y"))
df3$QtrYear <- sprintf("%s-%s", df3$Year, df3$Quarter)

```

# Data Preparation

## Data - Feature Engineering 

<span style="font-size: 18px;">
After bringing the stock.csv data, I created features to build a model.  
I additionally created `xxxx_SMA10` features which represent a Simple Moving Average.  
</span>


```{r echo=FALSE}
kable(head(df3), "html" ) %>%
  kable_styling((bootstrap_options = c("striped","hover")), font_size = 9, full_width = FALSE)
```


## Data - Sample Range 

<span style="font-size: 18px;">
I selected the year between $2011$ and $2012$ for this assignment.
</span>


```{r select_year, echo=FALSE}
qtrs <- df3 %>% 
  ungroup %>% 
  select(Year, QtrYear) %>% 
  unique() %>% 
  filter( between( Year, 2011, 2012) ) %>%
  arrange( QtrYear )

kable(qtrs, "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), font_size = 9, full_width = FALSE)
```


## Data - Train and Test 


<span style="font-size: 18px;">
As I selected year $2011$ and $2012$ I have five sets of training and testing.    
</span>


```{r echo=FALSE, fig.height=4}
# train/test are 3 and 1 qtrs in duration
trainLen <- 3
testLen <- 1

# rows in the data.frame
trainStart <- 1
trainStop <- nrow(qtrs) - trainLen

# lists we are collecting
trainSet <- list()
testSet <- list()

# identify and print the sets
while( trainStart <= trainStop ) {
  trainQ <- qtrs[seq(trainStart, length.out=trainLen), ]$QtrYear
  testQ <- qtrs[seq(trainStart+trainLen, length.out=testLen), ]$QtrYear
  
  print( "Set:" )
  print( trainQ ) 
  print( testQ )
  
  # extract the training ans testing from df3
  trainDF <- df3 %>% filter( QtrYear %in% trainQ )
  testDF <- df3 %>% filter( QtrYear %in% testQ )
  
  key <- sprintf("SET_%d", trainStart)
  
  trainSet[[key]] <- trainDF 
  testSet[[key]] <- testDF 
    
  trainStart <- trainStart + 1
}

```

```{r include=FALSE}
# iterating over the keys
train_dfs <- as.list(rep("", length(trainSet)))
test_dfs <-as.list(rep("", length(trainSet)))

for(i in seq_along(trainSet)) {
  aKey <-names(trainSet)[i]
  train_dfs[[i]] <- trainSet[[aKey]]
  test_dfs[[i]] <- testSet[[aKey]]
} 
```


## Data - Train and Test 


<span style="font-size: 18px;">
I am going to quickly check the number of rows in one train/test set.  
</span>

```{r sample_print, echo=FALSE}
print(sprintf("The number of cases in the first Train set: %s",nrow(train_dfs[[1]])))
print(sprintf("The number of cases in the first Test set: %s",nrow(test_dfs[[1]])))
```


# GLM

## Open_Close_PctChange 

<span style="font-size: 18px;">
I trained each 5 train set respectively, and combined the results in one single dataframe.  
I selected **outliers** as those which are larger than 5 standard deviations. 
I printed out each train set's RMSE and the number of outliers (combined).  
</span>

```{r include=FALSE}
h2o.init(nthreads=-1, max_mem_size='4G')
h2o.removeAll()
```

```{r include=FALSE}
rmse_list <- as.list(rep("", length(train_dfs)))
  
for (i in seq_along(train_dfs)){

  # get ready dataset
  train_test <-ResponseAndData_All("Open_Close_PctChange", train_dfs[[i]], test_dfs[[i]])
  
  # train the model
  results<- TrainAndPlot_3sd(train_test$pred, train_test$resp, 
                                train_test$train.hex, train_test$test.hex, train_test$test)
  
  # Show the result
  rmse_list[[i]] <- sprintf(" Model #%s RMSE: %s", i, results$sqrt_result )
  
  # Outlierss
  std <- sd(results$diff_df$diff)
  
  if (exists("glm_ocpc")){
    temp_ocpc = results$diff_df
    glm_ocpc = rbind(glm_ocpc, temp_ocpc)
    temp_outliers <- temp_ocpc %>%
      filter(abs(diff)>5*std)
    glm_ocpc_outliers=rbind(glm_ocpc_outliers, temp_outliers)
  }
  
  if (!exists("glm_ocpc")){
    glm_ocpc <- results$diff_df
    glm_ocpc_outliers <- glm_ocpc %>%
    filter(abs(diff)>5*std)
    
  }
}
```

```{r}
rmse_list

```

## Open_Close_PctChange 

<span style="font-size: 18px;">
I trained each 5 train set respectively, and combined the results in one single dataframe.  
I selected **outliers** as those which are larger than 5 standard deviations. 
I printed out each train set's RMSE and the number of outliers (combined).  
</span>

```{r}
nrow1 <- nrow(glm_ocpc_outliers)
print( sprintf("The number of outliers (larger than 5 STD) is %s", nrow1) )
```


## Open_Close_PctChange   


<span style="font-size: 18px;">
The below table shows the first $5$ rows of the detected outliers.  
</span>


```{r}
kable(head(glm_ocpc_outliers), "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), font_size = 9, full_width = FALSE)
```

```{r include=FALSE}
h2o.shutdown(prompt=TRUE)
```

## Daily_Return

<span style="font-size: 18px;">
I conducted the same modeling for the `Daily_Return` variable as well.  
Again, I printed RMSE and the number of outliers. 
</span>

```{r include=FALSE}
h2o.init(nthreads=-1, max_mem_size='4G')
h2o.removeAll()
```

```{r include=FALSE}
rmse_list <- as.list(rep("", length(train_dfs)))

for (i in seq_along(train_dfs)){

  # get ready dataset
  train_test <-ResponseAndData_All("Daily_Return", train_dfs[[i]], test_dfs[[i]])
  
  # train the model
  results<- TrainAndPlot_3sd(train_test$pred, train_test$resp, 
                                train_test$train.hex, train_test$test.hex, train_test$test)
  
  # Show the result
   rmse_list[[i]] <- sprintf(" Model #%s RMSE: %s", i, results$sqrt_result )
  
  # Outliers
  std <- sd(results$diff_df$diff)
  
  if (exists("glm_daily")){
   
    temp_daily = results$diff_df
    glm_daily = rbind(glm_daily, temp_daily)
    
    temp_outliers <- temp_daily %>%
      filter(abs(diff)>5*std)
    glm_daily_outliers=rbind(glm_daily_outliers, temp_outliers)
    
  }
  
  if (!exists("glm_daily")){
    
    glm_daily <- results$diff_df
    
    glm_daily_outliers <- glm_daily %>%
    filter(abs(diff)>5*std)
    
  }
}



```

```{r}
rmse_list

```

## Daily_Return

<span style="font-size: 18px;">
I conducted the same modeling for the `Daily_Return` variable as well.  
Again, I printed RMSE and the number of outliers. 
</span>

```{r}

nrow2 <- nrow(glm_daily_outliers)
print(sprintf("The number of outliers (larger than 5 STD) is %s", nrow2))
```


## Daily_Return

<span style="font-size: 18px;">
The below table shows the first 5 rows of the detected outliers.  
</span>

```{r echo=FALSE}
kable(head(glm_daily_outliers), "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), font_size = 9, full_width = FALSE)
```

```{r include=FALSE}
h2o.shutdown(prompt=TRUE)
```

# Random Forest

## Random Forest


<span style="font-size: 18px;">
  
  Although I am going to use the same target variables (`Open_Close_PctChange`, and `Daily_Return`)  
  for the random forest, I selected some of the original variables as predictive variables.  
  This is because non-valuable variables caused worse results for the random forests models.  
</span>


## Open_Close_PctChange 


<span style="font-size: 18px;">
As we can see from the below result, around $100$ outliers are detected in this case.  
</span>

```{r include=FALSE}
h2o.init(nthreads=-1, max_mem_size='4G')
h2o.removeAll()
```

```{r include=FALSE}
resp <- "Open_Close_PctChange"
pred <- c("Symbol", "DOW", "Sector",
          "Open_Mean10", "Close_Mean10", 
          "Open_Mean10_R", "Close_Mean10_R", 
          "Open_EMA10", "Close_EMA10",
          "Open_SMA10", "Close_SMA10")
  
for (i in seq_along(train_dfs)){

  # get ready dataset
  train_test <-Data_RF(train_dfs[[i]], test_dfs[[i]])

  # train the model
  results<- Train_Plot_RF(pred, resp, train_test$train.hex, train_test$test.hex, train_test$test)
  
  # Outliers
  
  std <- results$std
  
  if (exists("rf_ocpc")){
    
    temp_ocpc = results$diff_df
    rf_ocpc = rbind(rf_ocpc, temp_ocpc)
    
    temp_outliers <- temp_ocpc %>%
      filter(abs(diff)>5*std)
    rf_ocpc_outliers = rbind(rf_ocpc_outliers, temp_outliers)
  }
  
  if (!exists("rf_ocpc")){
    rf_ocpc <- results$diff_df
    rf_ocpc_outliers <- rf_ocpc %>%
    filter(abs(diff)>5*std)
    
  }
}


```

```{r}
nrow3 <- nrow(rf_ocpc_outliers)
print(sprintf("The number of outliers (larger than 5 STD) is %s",nrow3))
```


## Open_Close_PctChange

<span style="font-size: 18px;">
The below table shows detected outliers.  
</span>


```{r echo=FALSE}
kable(head(rf_ocpc_outliers), "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), font_size = 9, full_width = FALSE)
```

```{r include=FALSE}
h2o.shutdown(prompt=TRUE)
```


## Daily_Return

<span style="font-size: 16px;">
  Now, I will apply the same way to predict `Daily_Return`. 
  Below number of outliers were detected as we can see from the below resutls.
  In addition, the below table shows the first 5 rows of the detected outliers when the predictive variable is `Daily_Return`. 
</span>


```{r include=FALSE}
h2o.init(nthreads=-1, max_mem_size='4G')
h2o.removeAll()
``` 

```{r include=FALSE}

resp <- "Daily_Return"
pred <- c("Symbol", "DOW","Sector",
          "Open_Mean10", "Close_Mean10", 
          "Open_Mean10_R", "Close_Mean10_R", 
          "Open_EMA10", "Close_EMA10",
          "Open_SMA10", "Close_SMA10")
  
for (i in seq_along(train_dfs)){

  # get ready dataset
  train_test <-Data_RF(train_dfs[[i]], test_dfs[[i]])

  # train the model
  results<- Train_Plot_RF(pred, resp, train_test$train.hex, train_test$test.hex, train_test$test)
  
  # Outliers
  
  std <- results$std
  
  if (exists("rf_daily")){
    
    temp_daily = results$diff_df
    rf_daily = rbind(rf_daily, temp_daily)
    
    temp_outliers <- temp_daily %>%
      filter(abs(diff)>5*std)
    rf_daily_outliers = rbind(rf_daily_outliers, temp_outliers)
  }
  
  if (!exists("rf_daily")){
    rf_daily <- results$diff_df
    rf_daily_outliers <- rf_daily %>%
    filter(abs(diff)>5*std)
    
  }
}



```

```{r echo=FALSE}
print(sprintf("The number of outliers (larger than 5 STD) is %s",nrow(rf_daily_outliers)))

kable(head(rf_daily_outliers), "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), font_size = 9, full_width = FALSE)
```

```{r include=FALSE}
h2o.shutdown(prompt=FALSE)
```



# Business Insights

## Daily_Return Results


<span style="font-size: 18px;">
Since both showed pretty much similar results, I will show only `Daily_Return` here.  
I plotted the combined results from the two model (GLM and Random Forest) with the real values.  
</span>

```{r}
DailyReturn_df <- data.frame(glm_prediction = glm_daily$predictions, 
                           rf_prediction = rf_daily$predictions_rf,
                           real_value= glm_daily$Daily_Return)

DailyReturn_df <- DailyReturn_df %>%
  mutate(middle_prediction = (glm_prediction + rf_prediction)/2,
         indices=seq(nrow(DailyReturn_df))) %>%
  select(indices, real_value, glm_prediction, rf_prediction, middle_prediction)
  
DailyReturn_df %>%
  arrange(glm_prediction)%>%
  ggplot(aes(x=glm_prediction))+
  geom_point(aes(y=middle_prediction), color="red", alpha=0.5)+
  geom_point(aes(y=real_value), color="blue", alpha=0.5)+
  geom_line(aes(y=real_value), color="blue", linetype=2)+
  annotate(geom="text", x=3.5, y=5, label="Real Value\nline", color="blue")+
  annotate(geom="text", x=9, y=7.5, label="Predictions'\nCenter Points", color="red")+
  ggtitle("Daily_Return: Predictions VS. Real Values")
```


## Error

<span style="font-size: 16px;">
If we calculate the euclidean distance between   
 1. GLM Predictions VS. Real Values  
 2. RF Predictions VS. Real Values  
 3. Averaged Predictions VS. Real Values  
 As we can find from the below results, **glm** tends to predict more closely to the real values.  
</span>  


```{r echo=FALSE}
print( sprintf("Error with the GLM model: %s", dist(rbind(DailyReturn_df$glm_prediction, DailyReturn_df$real_value))[[1]]) )

print( sprintf ("Error with the Random Forest model: %s", dist(rbind(DailyReturn_df$rf_prediction, DailyReturn_df$real_value))[[1]]) )

print( sprintf ("Error with the Combined model: %s", dist(rbind(DailyReturn_df$middle_prediction, DailyReturn_df$real_value))[[1]]) )
```


## Final Insights


<span style="font-size: 16px;">
Finally, I inner joined the two outlier results (`glm_outliers` and `rf_outliers`) to find the commonly detected outliers.  
  Instead of looking at individual values, I decided to look at the `Sector` of the outliers, and look at the percentage of the outliers per each **sector**.  
This would be interesting information for investors who would like to invest stocks based on their *risk sensitivity*. 

</span>  


```{r}
daily_outliers_result <- glm_daily_outliers %>% inner_join(rf_daily_outliers, by=c("sector","symbol","Daily_Return")) %>%
  select(date.x, sector, symbol, qrt_year.x, Daily_Return)

daily_group <- glm_daily %>% 
  group_by(sector) %>%
  summarise(num_count=n())

buz_daily <-daily_outliers_result %>%
  group_by(sector) %>%
  summarise(num_count=n()) %>%
  right_join(daily_group, by="sector") %>%
  mutate(percentage= round( (num_count.x)/num_count.y *100 , 3) )

kable(buz_daily, "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), font_size = 9, full_width = FALSE)%>%
  row_spec(7, bold = T, color = "white", background = "skyblue")
```
