---
title: "Assignment 4, Revised Version"
author: "Yerin Jung"
date: "6/23/2018"
output:   
  html_document:
    theme: lumen
    number_sections: yes
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r functions, echo=FALSE}
source("health-functions.R")
```


# Data 


```{r data, echo=FALSE}
#setwd("/Users/Jung-yerin/Desktop/SUMMER_2018/AD/AnomalyDetection/week4")
payment <- read_csv("inpatientCharges.csv")
```

  The data frame contains information regarding a national summary of inpatient charge data by Medicare severity diagnosis related group (MS-DRG), originally from [Data.CMS](https://data.cms.gov/Medicare-Inpatient/National-Summary-of-Inpatient-Charge-Data-by-Medic/efwk-h4x3).  
  

# Table 
 
 <br />
 The following table shows the first 4 rows of the data frame.  
 I changed the column names in a more convenient way to analyze.  
  
  <br />
  
  
  
```{r echo=FALSE}
# Changing column names to remove space
names(payment) <- c("DRG_Def", "Pro_Id", "Pro_Name","Pro_Add","Pro_City","Pro_State","Pro_Zip",
                    "HRR","Tot_Discharges","Avg_Cov_Charges","Avg_Tot_Payments","Avg_Med_Payments")

kable(head(payment,4),"html")%>%
  kable_styling((bootstrap_options = c("striped","hover")), font_size = 10) %>%
  scroll_box(height = "100%", width = "90%")
```
  
  <br />
  
  The table below is a summary table of our dataset.  
  We will walk through each column by column in the following parts.  
  
  
  <br />
  
  
  
```{r echo=FALSE}
kable(summary(payment), "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), font_size = 10) %>%
  scroll_box(height = "100%", width = "90%")
```
  
  <br />
  
  First, `DRG_Def` represents classification system that groups similar clinical conditions (diagnoses) and  
  the procedures furnished by the hospital during the stay. [Data.CMS](https://data.cms.gov/Medicare-Inpatient/National-Summary-of-Inpatient-Charge-Data-by-Medic/efwk-h4x3)  
  
  
```{r echo=FALSE}
kable(head(payment,2),"html") %>%
  kable_styling(full_width = F, font_size = 10) %>%
  column_spec(1, bold = T, border_right = T, width = "100%", background = "skyblue")%>%
  scroll_box(height = "100%", width = "90%")
```
  
  <br />
  
  Second, `Pro_Id` stands for the id number of each provider(hospital).  
  This can be a key value for further analyses.  
  
  <br />
  
  
```{r echo=FALSE}
kable(head(payment,2),"html") %>%
  kable_styling(full_width = F, font_size = 10) %>%
  column_spec(2, bold = T, border_right = T, width = "100%", background = "skyblue") %>%
  scroll_box(height = "100%", width = "90%")
```
 
  <br />
  
  Third, `Pro_Name`, `Pro_Add`, `Pro_City`, `Pro_State`, `Pro_Zip`, and `HRR` represent information of the providers.  
  
  <br />
  
  
```{r echo=FALSE}
kable(head(payment,2),"html") %>%
  kable_styling(full_width = F, font_size = 10) %>%
  column_spec(c(3,4,5,6,7,8), bold = T, border_right = T, width = "100%", background = "skyblue")%>%
  scroll_box(height = "100%", width = "90%")
```
  
  
  <br />
  
  Fourth, `Tot_Discharges` represents The number of discharges billed by all providers for inpatient hospital services. [Data.CMS](https://data.cms.gov/Medicare-Inpatient/National-Summary-of-Inpatient-Charge-Data-by-Medic/efwk-h4x3)  
  
  <br />
  
  
```{r echo=FALSE}
kable(head(payment,2),"html") %>%
  kable_styling(full_width = F, font_size = 10) %>%
  column_spec(9, bold = T, border_right = T, width = "100%", background = "skyblue")%>%
  scroll_box(height = "100%", width = "90%")
```
  
  <br />
  
  Fifth, `Avg_Cov_Charges` represents The average charge of all provider's services covered by Medicare for discharges in the DRG.  
  These will vary from hospital to hospital because of differences in hospital charge structures. [Data.CMS](https://data.cms.gov/Medicare-Inpatient/National-Summary-of-Inpatient-Charge-Data-by-Medic/efwk-h4x3)  
  
  <br />
  
  
```{r echo=FALSE}
kable(head(payment,2),"html") %>%
  kable_styling(full_width = F, font_size = 10) %>%
  column_spec(10, bold = T, border_right = T, width = "100%", background = "skyblue")%>%
  scroll_box(height = "100%", width = "90%")
```
  
  <br />
  
  Sixth, `Avg_Tot_Payments` shows The average total payments to all providers for the DRG including the MS-DRG amount, teaching,  disproportionate share, capital, and outlier payments for all cases. [Data.CMS](https://data.cms.gov/Medicare-Inpatient/National-Summary-of-Inpatient-Charge-Data-by-Medic/efwk-h4x3)  
  
  <br />
  

```{r echo=FALSE}
kable(head(payment,2),"html") %>%
  kable_styling(full_width = F, font_size = 10) %>%
  column_spec(11, bold = T, border_right = T, width = "100%", background = "skyblue")%>%
  scroll_box(height = "100%", width = "90%")
```
  
  <br />
  
  Finally, `Avg_Med_Payment` represents The average amount that Medicare pays to the provider for Medicare's share of the MS-DRG.  
  [Data.CMS](https://data.cms.gov/Medicare-Inpatient/National-Summary-of-Inpatient-Charge-Data-by-Medic/efwk-h4x3)  
  
  <br />
  
  
```{r echo=FALSE}
kable(head(payment,2),"html") %>%
  kable_styling(full_width = F, font_size = 10) %>%
  column_spec(12, bold = T, border_right = T, width = "100%", background = "skyblue")%>%
  scroll_box(height = "100%", width = "90%")
```



# Columns

  <br />
 
 First of all, I will parse the numbers from Drug definition to see graphs effectively  
 
 
```{r}
payment <- parse_num(payment)
```
  
  <br />
  
  I decided to examine four quantitative comluns(`Tot_Discharges`, `Avg_Cov_Charges`, `Avg_Tot_Payments`, and `Avg_Med_Payments`) and one qualitative column(`Pro_State`).  
  
  <br />
  
## Tot_Discharges

```{r echo=FALSE}
payment %>%
  filter(Tot_Discharges > 800)%>%
  ggplot(aes(x=Tot_Discharges, fill=DRG_Num))+
  geom_histogram()+
  ggtitle("Histogram of Total Discharges over 800 Times")
  #theme(legend.position="none")
```
  
  <br />
  
  As we can see from the above chart, drug number $470$ and $871$ tended to be discharged more than the other drug definitions.  
  The information about the two drug-definitions are as follows:  
  
  - 470 : MAJOR JOINT REPLACEMENT OR REATTACHMENT OF LOWER EXTREMITY W/O MCC
  - 871 : SEPTICEMIA OR SEVERE SEPSIS W/O MV 96+ HOURS W MCC  
  
  <br />
  
## Avg_Cov_Charges, Avg_Tot_Payments, and Avg_Med_Payments

<br />

```{r echo=FALSE}
payment %>%
  ggplot(aes(x=Avg_Cov_Charges))+
  geom_histogram(binwidth = 1000, fill="skyblue")+
  geom_vline(xintercept =63474, linetype="dashed",color="red")+
  annotate(geom="text", x=250000, y=1000, label="3rd Quartile")+
  ggtitle("Histogram of Average Covered Charges with 3rd Quartile")
```

<br />

```{r echo=FALSE}
payment %>%
  ggplot(aes(x=Avg_Tot_Payments))+
  geom_histogram(binwidth = 1000, fill="skyblue")+
  geom_vline(xintercept =14655, linetype="dashed",color="red")+
  annotate(geom="text", x=50000, y=1000, label="3rd Quartile")+
  ggtitle("Histogram of Average Total Payments with 3rd Quartile")
```

<br />


```{r echo=FALSE}
payment %>%
  ggplot(aes(x=Avg_Med_Payments))+
  geom_histogram(binwidth = 1000, fill="skyblue")+
  geom_vline(xintercept =12442.5, linetype="dashed",color="red")+
  annotate(geom="text", x=40000, y=1000, label="3rd Quartile")+
  ggtitle("Histogram of Average Medicare Payments with 3rd Quartile")
```

<br />
  
  As we can see from the above three charts, all the variables are highly right-skewed.  
  Since all the drug definitions have different charges, we might need to group them by drug group and state as well.  
  
  <br />
  
## Pro_State

<br />


```{r echo=FALSE}
payment %>%
  group_by(Pro_State)%>%
  mutate(high_state =as.factor( ifelse(n()<10000,0,1) ))%>%
  ggplot(aes(x=Pro_State, fill=high_state)) +
  geom_bar(stat="count")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position="none")+
  ggtitle(label="Histogram of Count of Rows by State", subtitle = "Blue Colored is over 10,000 Counts")
```
<br />

  In the above chart, the states that have more than $10,000$ counts were colored by blue.  
  I assume that they have more counts of rows just because they are the top four states in population ranking.  
  
<br />

  
# New Features

## Comparing Standardized Total Payment 

<br />


### Lead

<br />

  Generally, total payment of a care depends on the **region(state)** and type of **drug function**.  
As such, to see whether there is abnormally high total payment (`Avg_Tot_Payments`), I decided to standardize the value, using the mean and standard deviation of `Avg_Tot_Payments` grouped by Drug Definition (`Drug_Num`) and Provider State (`Pro_State`).  

   <br />
  
  
### EDA

<br />

  To do so, first, I should count the number of providers in each drug definition group.  
  If there are just a few numbers of providers in the group, it might be less rational to compare the total payments.  
  (In many cases there was only one provider in a drug function group.) 

<br />

```{r}
# Making groups dependin on the number of providers in each drug group
payment1 <- group_class(payment)
```


<br />

  Here, as I mentioned, I divided the dataset into three groups: `1_to_945`, `946_to_1890`, and `1890_to_2837`:  
  - `1_to_945`: The number of entities in a drug definition group (`DRG_Def`) is a range of 1 to 946   
  - `946_to_1890`: The number of entities in a drug definition group (`DRG_Def`) is a range of 946 to 1890  
  - `1890_to_2837`: The number of entities in a drug definition group (`DRG_Def`) is a range of 1890 to 2837  

<br />

```{r echo=FALSE}
payment1_tail <- payment1 %>%
  filter(n_row_group==1)%>%
  select(DRG_Def,n_row_group, group_class) %>%
  tail()

kable(payment1_tail, "html") %>%
  kable_styling((bootstrap_options = c("striped","hover")), full_width = FALSE, font_size=10)%>%
  column_spec(column=2, bold = T, border_right = T, width = "2Oem", background = "skyblue") %>%
  scroll_box(height = "100%", width = "90%")
```

<br />

  As we can see from the above table, there are certain drug definition groups that have only one entity(case).  
  
  <br />
  <br />
  
  Then, I plotted `Avg_Tot_Payments` from only `1890_to_2837` group, to see the general trends as follows. 
  
  <br />
  
```{r echo=FALSE}
# The graph of the third group, which has relatively many providers in each drug group
payment1 %>%
  filter(group_class=="1890_to_2837")%>%
  ggplot(aes(x=Avg_Tot_Payments, y=DRG_Num))+
  geom_point(colour="skyblue")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  geom_vline(xintercept = 50000, linetype="dashed", color="red")+
  ggtitle("Avg Total Payments of the Drugs That Has More Than 1890 Cases within a Group")
```

  <br />
   
  It looks like we see some outliers, though, we should not be so sure, since the price can also vary from states.  
  
  <br />
  
```{r}
# Making Three Columns after grouping by DRG_Num and Pro_State
payment1<-add_standard_payment(payment1)
```


<br />
  
  I generated three columns as follows:    
  - `Mu_DRG_State_Total_Payment`: Mean of `Avg_Tot_Payments` grouped by drug group (`Drug_Num`) and state (`Prov_State`).  
  - `Sd_DRG_State_Total_Payment`: Standard deviation of `Avg_Tot_Payments` grouped by drug group (`Drug_Num`) and state (`Prov_State`).  
  - `Std_DRG_State_Total_Payment`: Standardized `Avg_Tot_Payments` using Z-score with `Mu_DRG_State_Total_Payment` and `Sd_DRG_State_Total_Payment`  
  
  

$$Z=\frac{X-\mu }{\sigma }=\frac{AvgTotalPayments-Mu(AvgTotalPayments)}{Sd(AvgTotalPayments)}$$

<br />

```{r echo=FALSE}
# Viewing some examples
payment1_exp<- payment1 %>%
  select(DRG_Num,Pro_State,Mu_DRG_State_Total_Payment,Sd_DRG_State_Total_Payment,Std_DRG_State_Total_Payment) %>%
  filter(DRG_Num=="310")%>%
  head()

kable(payment1_exp,"html")%>%
  kable_styling((bootstrap_options = c("striped","hover")), full_width = FALSE, font_size=10) %>%
  scroll_box(height = "100%", width = "90%")
```

<br />

## Conclusion

<br />

  After dividing into three group depending on the number of providers in drug_definition,    
  I filtered only the second and third group, which basically have more than 946 providers in each drug definition group.  

<br />

```{r echo=FALSE}
# Plotting standardized total_payment
payment1 %>%
  filter(group_class=="1890_to_2837"|group_class=="946_to_1890" ) %>%
  ggplot(aes(x=Pro_Id, y=Std_DRG_State_Total_Payment))+
  geom_point(alpha=0.5, color="skyblue")+
  theme(axis.text.x = element_blank())+
  geom_hline(yintercept = 3, linetype="dashed",color="red")+
  annotate(geom="text", x=70000, y=4, label="+3 Standard Deviation")+
  geom_hline(yintercept = 7, linetype="dashed",color="red")+
  annotate(geom="text", x=70000, y=8, label="+7 Standard Deviation")+
  ggtitle("Standardized Total Payment by State and Drug Group")
```

<br />

  As we can see from the above graph, we can notice a few distinctively high standardized values, comparing to the other providers. We might want to further exam those values.  
  Since they were outliers even after they were standardized by the mean grouped by state and drug as well, I believe that it has room to be a fraud case.  

<br />


## Standardized Difference Between Medical Coverage and Medical Payment

### Lead
  
  
  <br />

  Since `Average Covered Charges` would vary from hospital to hospital, we should not simply assume a fraud, just because a provider overly charged Medicare to pay than they were supposed to cover.  
  However, we can calculate the mean difference of `Average Covered Charges` and `Average Medicare Payments` per each Drug group, and see whether an entity has significant difference than the other in the same group. 
  
<br />

### EDA


<br />

```{r}
# Making Four New Columns
payment2<-add_standard(payment1)
```
  
  <br />
  
  First, I made three columns `Diff_MedCov_MedPay`, `Mu_Diff`, `Sd_Diff`, and `Standard_Diff`.  
  - `Diff_MedCov_MedPay`: The difference between the amount that is covered by Medicare(`Avg_Cov_Charges`) and the amount Medicare actually paid (`Avg_Med_Payments`)  
  - `Mu_Diff`: Average of `Diff_MedCov_MedPay` grouped by drug definition(`Drug_Num`). I grouped by drug definition, since some drugs might just require more money than it is generally covered.  
  - `Sd_Diff`: Standard deviation of `Diff_MedCov_MedPay`.  
  - `Standard_Diff`: Standardized values of `Diff_MedCov_MedPay`, using Z-score formula with `Mu_Diff` and `Sd_Diff`  
  
  

$$Z=\frac{X-\mu }{\sigma }=\frac{DiffMedCovMedPay-MuDiff}{Sd_Diff}$$
  
  <br />
  
  This table below shows the examples of how standardized difference looks like. 
  
  <br />
  
```{r echo=FALSE}
payment2_exp<- payment2 %>%
  select(DRG_Num, Diff_MedCov_MedPay, Mu_Diff, Sd_Diff, Standard_Diff ) %>%
  head()

kable(payment2_exp, "html")%>%
  kable_styling((bootstrap_options = c("striped","hover")), full_width = FALSE, font_size=10) %>%
  column_spec(5, bold = T, border_right = T, width = "20em", background = "skyblue")
```

<br />


  As we can see from the above table, most of the providers were centered around zero. However, there were several providers with the standardized difference even less than -2.  
  This means that those providers were paid way more than they normally have been covered by Medicare.  
   
  <br />
  

```{r echo=FALSE}
# Plotting a histogram
payment2 %>%
  ggplot(aes(x=Standard_Diff))+
  geom_histogram(fill="skyblue")+
  ggtitle("Standardized Difference between Avg_Cov_Charges and Avg_Med_Payments")+
  xlab(label="Standardized Difference")+
  geom_vline(xintercept = -2, linetype="dashed")+
  annotate(geom="text", x=-3.8, y=40000, label="Area where \nMedicare paid \n way more than \nthey mostly cover")
```
  
  
  

  
  
## Conclusion
  
  <br />
  
   As such, I looked at the providers with the standardized difference even less than -2.  
   
   <br />
   
```{r echo=FALSE}
# Filtering intresting data
payment2_interest <-payment2 %>% 
  select(DRG_Num, Pro_Id, Avg_Cov_Charges,Avg_Med_Payments,Standard_Diff)%>%
  filter(Standard_Diff < -2)

kable(payment2_interest,"html")%>%
  kable_styling((bootstrap_options = c("striped","hover")), font_size=10) %>%
  column_spec(column=c(2,5), bold = T, border_right = T, width = "20em", background = "skyblue")%>%
  row_spec(11, bold = T, color = "white", background = "green")
```

   <br />

  As we can see from the above table, there were $17$ providers who have a standardized difference of less than -2.  
  This means the amount of money that Medicare paid to the providers was significantly higher than the amount to the other providers in the same drug definition.  
  One of them has even less than -4, as we can see from the green line.  
  We may assume that those cases have a possibility to be a fraud, and examine them.  
  
  <br />


## Number of Discharges

### Lead
  
  
  <br />
  
  I would like to examine the number of discharges by drug and state, to detect unbundling.  
  I believe that this feature can be powerful to detect anomalies when it is combined with the previously defined columns.  
  
  <br />
  
### EDA
  
  <br />
  
  First, I made `Mean_Discharge` and `Stdv_Discharges` to eventually create `Standardized_Discharges` column.  
  `Standardized_Discharges` column represents the standardized value of the number of discharges, grouped by drug and state.  
  
  <br />
  
```{r}
payment3 <- add_standard2(payment2)
```

  <br />
  
  We can check the standardized values from the below chart.  
  
  <br />
  
```{r echo=FALSE}
payment3_exp <-payment3 %>%
  select(DRG_Num, Pro_Id, Tot_Discharges, Standardized_Discharges)%>%
  tail(100)%>%head()

kable(payment3_exp,"html")%>%
  kable_styling(full_width = FALSE, font_size = 10)%>%
  column_spec(4, bold = T, border_right = T, width = "20em", background = "skyblue")
```

  
  <br />
  
  Next, I looked at the histogram of the standardized number of discharges.   
  We can see that they are right-skewed.  
  

  <br />
  
```{r echo=FALSE}
payment3 %>%
  ggplot(aes(x=Standardized_Discharges, na.rm=TRUE))+
  geom_histogram(fill="skyblue")+
  geom_vline(xintercept = 2, linetype="dashed",color="red")+
  annotate(geom="text", x=3.5, y=30000, label="+2 Standard Deviation")+
  ggtitle("Histogram of Standardized Number of Discharges")
  
```

  <br />
  
  
  Finally, I combined the standardized number of discharges and the standardized difference. (`Standardized_Discharges` and`Standard_Diff`) and plotted as follows.  
  
  <br />
  
```{r echo=FALSE}
payment3 %>%
  mutate(Col=as.factor(ifelse(Standard_Diff > 5, 1,0)))%>%
  filter(Standardized_Discharges>2)%>%
  ggplot(aes(x=Pro_Id, y=Standard_Diff,color=Col, na.rm=TRUE))+
  geom_point(alpha=0.5)+
  theme(legend.position="none")+
  ylab(label="Diff Btwn the Madicare Coverage and Payment")+
  ggtitle(label = "Providers With Number of Dischargs Twice Than the Average",
          subtitle = "Blue colored are the ones who have bigger difference between coverage and paid")
```
  <br />
  
  As we can see from the above graph, there were several providers with high `Standardized_Discharges`, and high `Standard_Diff` as well.  
  (Those points have more than 2 `Standardized_Discharges`)
  
  <br />
  

### Conclusion
  
  <br />
  
  Since those $11$ cases are where the difference between Medicare coverage and Medicare payment is huge, and where more discharges happened as well, we might want to further examine those anomalies.  
  
  <br />
  
```{r echo=FALSE}
payment3_inter<- payment3 %>%
  filter(Standardized_Discharges>2)%>%
  filter(Standard_Diff>=5)%>%
  select(DRG_Num, Pro_Id, Standardized_Discharges, Standard_Diff)

kable(payment3_inter,"html")%>%
  kable_styling(full_width = FALSE, font_size = 10)%>%
  column_spec(c(3,4), bold = T, border_right = T, width = "20em", background = "skyblue")
```

  
  
  
  <br />
  <br />
  <br />
  <br />
  <br />
  











